---
Title: "Taxa-based Models: Arcsine Square Root Transformation, Forward Stepwise Regression"
Author: Henry Paz (henry.paz@huskers.unl.edu)
Output:
  html_document:
    keep_md: yes
---

The following arcsine square root transforms the summarized test sets at family level and performs forward stepwise regressions of taxa-based models for average daily feed intake (ADFI), average daily gain (ADG), feed efficiency (ADG/ADFI) within heifer and steer cohort.

## Summarized test sets at family level for heifer and steer cohorts

```{r, engine='bash'}
sed 's/#OTU ID/taxonomy/g' biom_files/otu_table_test_heifer_L5.txt > r_inputs/otu_table_test_heifer_L5.txt

sed 's/#OTU ID/taxonomy/g' biom_files/otu_table_test_steer_L5.txt > r_inputs/otu_table_test_steer_L5.txt
```

```{r}
#Load packages
library(stringr)

#heifer data
heifer_test_family <- read.table("r_inputs/otu_table_test_heifer_L5.txt", sep="\t", header=T)
samples_heifer <- heifer_test_family[,-1]
taxa_heifer <- as.data.frame(str_split_fixed(heifer_test_family$taxonomy, ";", 5))
names(taxa_heifer)[5] <- "taxonomy"
taxa_heifer$taxonomy <- sub("f__", "", taxa_heifer$taxonomy)
taxa_heifer$taxonomy <- sub("\\]", "", taxa_heifer$taxonomy)
taxa_heifer$taxonomy <- sub("\\[", "", taxa_heifer$taxonomy)
taxa_heifer$taxonomy <- sub("^$", "No Assigned Family", taxa_heifer$taxonomy)
taxa_heifer$taxonomy <- sub("Other", "No Assigned Family", taxa_heifer$taxonomy)
taxa_heifer$taxonomy[1] <- "NoAssignedFamily1"
taxa_heifer$taxonomy[60] <- "NoAssignedFamily2"
taxa_heifer$taxonomy[61] <- "NoAssignedFamily3"
taxa_heifer$taxonomy[114] <- "NoAssignedFamily4"
taxa_heifer$taxonomy[134] <- "NoAssignedFamily5"
taxa_heifer$taxonomy[135] <- "NoAssignedFamily6"
taxa_heifer$taxonomy[165] <- "NoAssignedFamily7"
taxa_heifer$taxonomy[183] <- "NoAssignedFamily8"
taxa_heifer$taxonomy[188] <- "NoAssignedFamily9"
taxa_heifer$taxonomy[223] <- "NoAssignedFamily10"
taxa_heifer$taxonomy[228] <- "NoAssignedFamily11"
taxa_heifer$taxonomy[241] <- "NoAssignedFamily12"
taxa_heifer$taxonomy[264] <- "NoAssignedFamily13"
taxa_heifer$taxonomy[265] <- "NoAssignedFamily14"
taxa_heifer$taxonomy[284] <- "NoAssignedFamily15"
taxa_heifer$taxonomy[286] <- "NoAssignedFamily16"
taxa_heifer$taxonomy[289] <- "NoAssignedFamily17"
taxa_heifer$taxonomy[290] <- "NoAssignedFamily18"
taxa_heifer$taxonomy[291] <- "NoAssignedFamily19"
taxa_heifer$taxonomy[306] <- "NoAssignedFamily20"
taxa_heifer$taxonomy[167] <- "R4_45B"
#validation set
taxa_heifer$taxonomy[71] <- "S24_7"
taxa_heifer$taxonomy[75] <- "p_2534_18B5"
heifer_test_fam2 <- merge(samples_heifer, taxa_heifer, by="row.names")
heifer_test_fam3 <- heifer_test_fam2[,-c(1,111,112,113,114)]

write.table(heifer_test_fam3, sep="\t", file="r_inputs/heifer_test_family_final.txt", row.names=F, col.names=T, quote=F)

#steer data
steer_test_family <- read.table("r_inputs/otu_table_test_steer_L5.txt", sep="\t", header=T)
samples_steer <- steer_test_family[,-1]
taxa_steer <- as.data.frame(str_split_fixed(steer_test_family$taxonomy, ";", 5))
names(taxa_steer)[5] <- "taxonomy"
taxa_steer$taxonomy <- sub("f__", "", taxa_steer$taxonomy)
taxa_steer$taxonomy <- sub("\\]", "", taxa_steer$taxonomy)
taxa_steer$taxonomy <- sub("\\[", "", taxa_steer$taxonomy)
taxa_steer$taxonomy <- sub("^$", "No Assigned Family", taxa_steer$taxonomy)
taxa_steer$taxonomy <- sub("Other", "No Assigned Family", taxa_steer$taxonomy)
taxa_steer$taxonomy[1] <- "NoAssignedFamily1"
taxa_steer$taxonomy[60] <- "NoAssignedFamily2"
taxa_steer$taxonomy[61] <- "NoAssignedFamily3"
taxa_steer$taxonomy[114] <- "NoAssignedFamily4"
taxa_steer$taxonomy[134] <- "NoAssignedFamily5"
taxa_steer$taxonomy[135] <- "NoAssignedFamily6"
taxa_steer$taxonomy[165] <- "NoAssignedFamily7"
taxa_steer$taxonomy[183] <- "NoAssignedFamily8"
taxa_steer$taxonomy[188] <- "NoAssignedFamily9"
taxa_steer$taxonomy[223] <- "NoAssignedFamily10"
taxa_steer$taxonomy[228] <- "NoAssignedFamily11"
taxa_steer$taxonomy[241] <- "NoAssignedFamily12"
taxa_steer$taxonomy[264] <- "NoAssignedFamily13"
taxa_steer$taxonomy[265] <- "NoAssignedFamily14"
taxa_steer$taxonomy[284] <- "NoAssignedFamily15"
taxa_steer$taxonomy[286] <- "NoAssignedFamily16"
taxa_steer$taxonomy[288] <- "NoAssignedFamily17"
taxa_steer$taxonomy[290] <- "NoAssignedFamily18"
taxa_steer$taxonomy[291] <- "NoAssignedFamily19"
taxa_steer$taxonomy[306] <- "NoAssignedFamily20"
taxa_steer$taxonomy[71] <- "S24_7"
taxa_steer$taxonomy[75] <- "p_2534_18B5"
#validation set
taxa_steer$taxonomy[167] <- "R4_45B"
steer_test_fam2 <- merge(samples_steer, taxa_steer, by="row.names")
steer_test_fam3 <- steer_test_fam2[,-c(1,108,109,110,111)]

write.table(steer_test_fam3, sep="\t", file="r_inputs/steer_test_family_final.txt", row.names=F, col.names=T, quote=F)
```

## Arcsine square root transformation of test sets at family level for heifer and steer cohorts

```{r}
#Load packages
library(dplyr)

#Differential families sets
differential_fam_heifer <- read.table("differential_otus/heifer_fam_differential.txt", sep="\t", header=F)
differential_fam_steer <- read.table("differential_otus/steer_fam_differential.txt", sep="\t", header=F)

#heifer data
heifer_family <- read.table("r_inputs/heifer_test_family_final.txt", sep="\t", header=T, fill=T, stringsAsFactors=F)
heifer_taxa <- subset(heifer_family, select = c(taxonomy))
heifer_data <- heifer_family[,-110]
#arcsine square root transformation
trans.arcsine <- function(x){asin(sign(x) * sqrt(abs(x)))}
heifer_data[, 1:109] <- as.data.frame(lapply(heifer_data[, 1:109], FUN = function(x) {sapply(x, FUN = trans.arcsine)}))
merged_heifer <- merge(heifer_taxa, heifer_data, by = "row.names")
merged_heifer <- merged_heifer[,-1]
heifer_filtered <- merged_heifer[which(merged_heifer$taxonomy %in% differential_fam_heifer$V1),]
heifer_trans <- as.data.frame(t(heifer_filtered), stringsAsFactors=F)
colnames(heifer_trans) = heifer_trans[1, ]
heifer_trans = heifer_trans[-1, ]
heifer_trans_final <- add_rownames(heifer_trans, "SampleID")
write.table(heifer_trans_final, file="stepwise_reg/heifer_fam_final.txt", sep="\t", col.names=T, row.names=F, quote=F)

#steer data
steer_family <- read.table("r_inputs/steer_test_family_final.txt", sep="\t", header=T, fill=T, stringsAsFactors=F)
steer_taxa <- subset(steer_family, select = c(taxonomy))
steer_data <- steer_family[,-107]
#arcsine square root transformation
trans.arcsine <- function(x){asin(sign(x) * sqrt(abs(x)))}
steer_data[, 1:106] <- as.data.frame(lapply(steer_data[, 1:106], FUN = function(x) {sapply(x, FUN = trans.arcsine)}))
merged_steer <- merge(steer_taxa, steer_data, by = "row.names")
merged_steer <- merged_steer[,-1]
steer_filtered <- merged_steer[which(merged_steer$taxonomy %in% differential_fam_steer$V1),]
steer_trans <- as.data.frame(t(steer_filtered), stringsAsFactors=F)
colnames(steer_trans) = steer_trans[1, ]
steer_trans = steer_trans[-1, ]
steer_trans_final <- add_rownames(steer_trans, "SampleID")
write.table(steer_trans_final, file="stepwise_reg/steer_fam_final.txt", sep="\t", col.names=T, row.names=F, quote=F)

#Heifer validation set
validation_heifer <- merged_steer[which(merged_steer$taxonomy %in% differential_fam_heifer$V1),]
validation_heifer_trans <- as.data.frame(t(validation_heifer), stringsAsFactors=F)
colnames(validation_heifer_trans) = validation_heifer_trans[1, ]
validation_heifer_trans = validation_heifer_trans[-1, ]
validation_heifer_final <- add_rownames(validation_heifer_trans, "SampleID")
write.table(validation_heifer_final, file = "stepwise_reg/validation_heifer_fam.txt", sep="\t", col.names=T, row.names=F, quote=F)

#Steer validation set
validation_steer <- merged_heifer[which(merged_heifer$taxonomy %in% differential_fam_steer$V1),]
validation_steer_trans <- as.data.frame(t(validation_steer), stringsAsFactors=F)
colnames(validation_steer_trans) = validation_steer_trans[1, ]
validation_steer_trans = validation_steer_trans[-1, ]
validation_steer_final <- add_rownames(validation_steer_trans, "SampleID")
write.table(validation_steer_final, file = "stepwise_reg/validation_steer_fam.txt", sep="\t", col.names=T, row.names=F, quote=F)
```

#Forward stepwise regression of taxa-based models at family level for average daily feed intake (ADFI), average daily gain (ADG), feed efficiency (ADG/ADFI) within heifer and steer cohort

```{r}
#Load packages
library(car)

#Heifer data 
mapping_heifer <- read.table("r_inputs/mapping_test_heifer.txt", sep="\t", header=T)
row.names(mapping_heifer) <- mapping_heifer[,1]

#Heifer linear models for ADFI, ADG, and FE using breed composition as covariates 
lm_ADFI_heifer <- lm(ADFI ~ perANS +	perHHS + perARS +	perSHS + perBMS + perBRS + perBNS +	perSGS + perBVS +	perCHS + perCAS +	perGVS + perLMS +	perMAS + perSAS + perSMS + perHH + perAN + perSM + perCH + perM2 + perM3 + perRS + perRO, data=mapping_heifer)
ADFI_res_heifer <- as.data.frame(resid(lm_ADFI_heifer))
lm_ADG_heifer <- lm(ADG ~ perANS +	perHHS + perARS +	perSHS + perBMS + perBRS + perBNS +	perSGS + perBVS +	perCHS + perCAS +	perGVS + perLMS +	perMAS + perSAS + perSMS + perHH + perAN + perSM + perCH + perM2 + perM3 + perRS + perRO, data=mapping_heifer)
ADG_res_heifer <- as.data.frame(resid(lm_ADG_heifer))
lm_FE_heifer <- lm(FE ~ perANS +	perHHS + perARS +	perSHS + perBMS + perBRS + perBNS +	perSGS + perBVS +	perCHS + perCAS +	perGVS + perLMS +	perMAS + perSAS + perSMS + perHH + perAN + perSM + perCH + perM2 + perM3 + perRS + perRO, data=mapping_heifer)
FE_res_heifer <- as.data.frame(resid(lm_FE_heifer))

#Heifer dataframes
heifer_fam <- read.table("stepwise_reg/heifer_fam_final.txt", sep = "\t", header=T)
row.names(heifer_fam) <- heifer_fam[,1]
heifer_fam <- heifer_fam[,-1]
heifer_ADFI <- merge(ADFI_res_heifer,heifer_fam,by="row.names")
heifer_ADFI <- heifer_ADFI[,-1]
colnames(heifer_ADFI)[1] <- "ADFI"
heifer_ADG <- merge(ADG_res_heifer,heifer_fam,by="row.names")
heifer_ADG <- heifer_ADG[,-1]
colnames(heifer_ADG)[1] <- "ADG"
heifer_FE <- merge(FE_res_heifer,heifer_fam,by="row.names")
heifer_FE <- heifer_FE[,-1]
colnames(heifer_FE)[1] <- "FE"

#Heifer models
#stepwise regression forward ADFI
null_heifer_ADFI_fam=lm(ADFI~1, data=heifer_ADFI)
#summary(null_heifer_ADFI_fam)
full_heifer_ADFI_fam=lm(ADFI~., data=heifer_ADFI)
#summary(full_heifer_ADFI_fam)
step(null_heifer_ADFI_fam, scope=list(lower=null_heifer_ADFI_fam, upper=full_heifer_ADFI_fam), direction="forward")

#Model ADFI
model_heifer_ADFI_fam <- lm(formula = ADFI ~ NoAssignedFamily18 + Veillonellaceae + Coriobacteriaceae, data=heifer_ADFI)
summary(model_heifer_ADFI_fam)
##anova(model_heifer_ADFI_fam)
#Test multi-collinearity (variance inflation factor)
vif(model_heifer_ADFI_fam)
sqrt(vif(model_heifer_ADFI_fam)) > 2
#Homoscedasticity & Normality
layout(matrix(c(1,2,3,4),2,2))
plot(model_heifer_ADFI_fam)
#Extract observed and predicted values
OP_heifer_ADFI_fam <- data.frame(Observed = heifer_ADFI$ADFI, Predicted = fitted(model_heifer_ADFI_fam))
#Plot observed vs prediceted
par(mfrow=c(1,1))
plot(OP_heifer_ADFI_fam$Predicted, OP_heifer_ADFI_fam$Observed, xlab="Predicted Average Daily Feed Intake (kg/d)", ylab="Observed Average Daily Feed Intake (kg/d)")
OP=lm(Observed~Predicted, data=OP_heifer_ADFI_fam)
#summary(OP)
abline(OP)

#stepwise regression forward ADG
null_heifer_ADG_fam=lm(ADG~1, data=heifer_ADG)
#summary(null_heifer_ADG_fam)
full_heifer_ADG_fam=lm(ADG~., data=heifer_ADG)
#summary(full_heifer_ADG_fam)
step(null_heifer_ADG_fam, scope=list(lower=null_heifer_ADG_fam, upper=full_heifer_ADG_fam), direction="forward")

#Model ADG
model_heifer_ADG_fam <- lm(formula = ADG ~  Veillonellaceae + Coriobacteriaceae, data=heifer_ADG)
summary(model_heifer_ADG_fam)
#anova(model_heifer_ADG_fam)
#Test multi-collinearity (variance inflation factor)
vif(model_heifer_ADG_fam)
sqrt(vif(model_heifer_ADG_fam)) > 2
#Homoscedasticity & Normality
layout(matrix(c(1,2,3,4),2,2))
plot(model_heifer_ADG_fam)
#Extract observed and predicted values
OP_heifer_ADG_fam <- data.frame(Observed = heifer_ADG$ADG, Predicted = fitted(model_heifer_ADG_fam))
#Plot observed vs prediceted
par(mfrow=c(1,1))
plot(OP_heifer_ADG_fam$Predicted, OP_heifer_ADG_fam$Observed, xlab="Predicted Average Daily Gain (kg/d)", ylab="Observed Average Daily Gain (kg/d)")
OP=lm(Observed~Predicted, data=OP_heifer_ADG_fam)
#summary(OP)
abline(OP)

#stepwise regression forward FE
null_heifer_FE_fam=lm(FE~1, data=heifer_FE)
#summary(null_heifer_FE_fam)
full_heifer_FE_fam=lm(FE~., data=heifer_FE)
#summary(full_heifer_FE_fam)
step(null_heifer_FE_fam, scope=list(lower=null_heifer_FE_fam, upper=full_heifer_FE_fam), direction="forward")

#Model FE
model_heifer_FE_fam <- lm(formula = FE ~ Veillonellaceae + Coriobacteriaceae + NoAssignedFamily2, data=heifer_FE)
summary(model_heifer_FE_fam)
#anova(model_heifer_FE_fam)
#Test multi-collinearity (variance inflation factor)
vif(model_heifer_FE_fam)
sqrt(vif(model_heifer_FE_fam)) > 2
#Homoscedasticity & Normality
layout(matrix(c(1,2,3,4),2,2))
plot(model_heifer_FE_fam)
#Extract observed and predicted values
OP_heifer_FE_fam <- data.frame(Observed = heifer_FE$FE, Predicted = fitted(model_heifer_FE_fam))
#Plot observed vs prediceted
par(mfrow=c(1,1))
plot(OP_heifer_FE_fam$Predicted, OP_heifer_FE_fam$Observed, xlab="Predicted Feed Efficiency (ADG/ADFI)", ylab="Observed Feed Efficiency (ADG/ADFI)")
OP=lm(Observed~Predicted, data=OP_heifer_FE_fam)
##summary(OP)
abline(OP)

#Steer data 
mapping_steer <- read.table("r_inputs/mapping_test_steer.txt", sep="\t", header=T)
row.names(mapping_steer) <- mapping_steer[,1]

#Steer linear models for ADFI, ADG, and FE using breed composition as covariates 
lm_ADFI_steer <- lm(ADFI ~ perANS +	perHHS + perARS +	perSHS + perBMS + perBRS + perBNS +	perSGS + perBVS +	perCHS + perCAS +	perGVS + perLMS +	perMAS + perSAS + perSMS + perHH + perAN + perSM + perCH + perM2 + perM3 + perRS + perRO, data=mapping_steer)
ADFI_res_steer <- as.data.frame(resid(lm_ADFI_steer))
lm_ADG_steer <- lm(ADG ~ perANS +	perHHS + perARS +	perSHS + perBMS + perBRS + perBNS +	perSGS + perBVS +	perCHS + perCAS +	perGVS + perLMS +	perMAS + perSAS + perSMS + perHH + perAN + perSM + perCH + perM2 + perM3 + perRS + perRO, data=mapping_steer)
ADG_res_steer <- as.data.frame(resid(lm_ADG_steer))
lm_FE_steer <- lm(FE ~ perANS +	perHHS + perARS +	perSHS + perBMS + perBRS + perBNS +	perSGS + perBVS +	perCHS + perCAS +	perGVS + perLMS +	perMAS + perSAS + perSMS + perHH + perAN + perSM + perCH + perM2 + perM3 + perRS + perRO, data=mapping_steer)
FE_res_steer <- as.data.frame(resid(lm_FE_steer))

#steer dataframes
steer_fam <- read.table("stepwise_reg/steer_fam_final.txt", sep="\t", header=T)
row.names(steer_fam) <- steer_fam[,1]
steer_fam <- steer_fam[,-1]
steer_ADFI <- merge(ADFI_res_steer,steer_fam,by="row.names")
steer_ADFI <- steer_ADFI[,-1]
colnames(steer_ADFI)[1] <- "ADFI"
steer_ADG <- merge(ADG_res_steer,steer_fam,by="row.names")
steer_ADG <- steer_ADG[,-1]
colnames(steer_ADG)[1] <- "ADG"
steer_FE <- merge(FE_res_steer,steer_fam,by="row.names")
steer_FE <- steer_FE[,-1]
colnames(steer_FE)[1] <- "FE"

#steer models
#stepwise regression forward ADFI
null_steer_ADFI_fam=lm(ADFI~1, data=steer_ADFI)
#summary(null_steer_ADFI_fam)
full_steer_ADFI_fam=lm(ADFI~., data=steer_ADFI)
#summary(full_steer_ADFI_fam)
step(null_steer_ADFI_fam, scope=list(lower=null_steer_ADFI_fam, upper=full_steer_ADFI_fam), direction="forward")

#Model ADFI
model_steer_ADFI_fam <- lm(formula = ADFI ~ NoAssignedFamily13 + Paraprevotellaceae + Lachnospiraceae + Coriobacteriaceae + Bifidobacteriaceae + Erysipelotrichaceae, data=steer_ADFI)
summary(model_steer_ADFI_fam)
##anova(model_steer_ADFI_fam)
#Test multi-collinearity (variance inflation factor)
vif(model_steer_ADFI_fam)
sqrt(vif(model_steer_ADFI_fam)) > 2
#Homoscedasticity & Normality
layout(matrix(c(1,2,3,4),2,2))
plot(model_steer_ADFI_fam)
#Extract observed and predicted values
OP_steer_ADFI_fam <- data.frame(Observed = steer_ADFI$ADFI, Predicted = fitted(model_steer_ADFI_fam))
#Plot observed vs prediceted
par(mfrow=c(1,1))
plot(OP_steer_ADFI_fam$Predicted, OP_steer_ADFI_fam$Observed, xlab="Predicted Average Daily Feed Intake (kg/d)", ylab="Observed Average Daily Feed Intake (kg/d)")
OP=lm(Observed~Predicted, data=OP_steer_ADFI_fam)
#summary(OP)
abline(OP)

#stepwise regression forward ADG
null_steer_ADG_fam=lm(ADG~1, data=steer_ADG)
#summary(null_steer_ADG_fam)
full_steer_ADG_fam=lm(ADG~., data=steer_ADG)
#summary(full_steer_ADG_fam)
step(null_steer_ADG_fam, scope=list(lower=null_steer_ADG_fam, upper=full_steer_ADG_fam), direction="forward")

#Model ADG
model_steer_ADG_fam <- lm(formula = ADG ~ p_2534_18B5 + Lachnospiraceae + Coriobacteriaceae + Bifidobacteriaceae + NoAssignedFamily9, data=steer_ADG)
summary(model_steer_ADG_fam)
#anova(model_steer_ADG_fam)
#Test multi-collinearity (variance inflation factor)
vif(model_steer_ADG_fam)
sqrt(vif(model_steer_ADG_fam)) > 2
#Homoscedasticity & Normality
layout(matrix(c(1,2,3,4),2,2))
plot(model_steer_ADG_fam)
#Extract observed and predicted values
OP_steer_ADG_fam <- data.frame(Observed = steer_ADG$ADG, Predicted = fitted(model_steer_ADG_fam))
#Plot observed vs prediceted
par(mfrow=c(1,1))
plot(OP_steer_ADG_fam$Predicted, OP_steer_ADG_fam$Observed, xlab="Predicted Average Daily Gain (kg/d)", ylab="Observed Average Daily Gain (kg/d)")
OP=lm(Observed~Predicted, data=OP_steer_ADG_fam)
#summary(OP)
abline(OP)

#stepwise regression forward FE
null_steer_FE_fam=lm(FE~1, data=steer_FE)
#summary(null_steer_FE_fam)
full_steer_FE_fam=lm(FE~., data=steer_FE)
#summary(full_steer_FE_fam)
step(null_steer_FE_fam, scope=list(lower=null_steer_FE_fam, upper=full_steer_FE_fam), direction="forward")

#Model FE
model_steer_FE_fam <- lm(formula = FE ~ NoAssignedFamily13 + Paraprevotellaceae + NoAssignedFamily9, data=steer_FE)
summary(model_steer_FE_fam)
#anova(model_steer_FE_fam)
#Test multi-collinearity (variance inflation factor)
vif(model_steer_FE_fam)
sqrt(vif(model_steer_FE_fam)) > 2
#Homoscedasticity & Normality
layout(matrix(c(1,2,3,4),2,2))
plot(model_steer_FE_fam)
#Extract observed and predicted values
OP_steer_FE_fam <- data.frame(Observed = steer_FE$FE, Predicted = fitted(model_steer_FE_fam))
#Plot observed vs prediceted
par(mfrow=c(1,1))
plot(OP_steer_FE_fam$Predicted, OP_steer_FE_fam$Observed, xlab="Predicted Feed Efficiency (ADG/ADFI)", ylab="Observed Feed Efficiency (ADG/ADFI)")
OP=lm(Observed~Predicted, data=OP_steer_FE_fam)
#summary(OP)
abline(OP)

#Validation heifer data sets
validation_fam_heifer <- read.table("stepwise_reg/validation_heifer_fam.txt", sep="\t", header=T)
row.names(validation_fam_heifer) <- validation_fam_heifer[,1]
validation_fam_heifer <- validation_fam_heifer[,-1]
validation_heifer_fam_ADFI <- merge(ADFI_res_steer,validation_fam_heifer,by="row.names")
validation_heifer_fam_ADFI <- validation_heifer_fam_ADFI[,-1]
colnames(validation_heifer_fam_ADFI)[1] <- "ADFI"
validation_heifer_fam_ADG <- merge(ADG_res_steer,validation_fam_heifer,by="row.names")
validation_heifer_fam_ADG <- validation_heifer_fam_ADG[,-1]
colnames(validation_heifer_fam_ADG)[1] <- "ADG"
validation_heifer_fam_FE <- merge(FE_res_steer,validation_fam_heifer,by="row.names")
validation_heifer_fam_FE <- validation_heifer_fam_FE[,-1]
colnames(validation_heifer_fam_FE)[1] <- "FE"

#Validation heifer models
#Heifer ADFI
modelval_heifer_fam_ADFI <- lm(formula = ADFI ~ NoAssignedFamily18 + Veillonellaceae + Coriobacteriaceae, data=validation_heifer_fam_ADFI)
summary(modelval_heifer_fam_ADFI)

#Heifer ADG
modelval_heifer_fam_ADG <- lm(formula = ADG ~ Veillonellaceae + Coriobacteriaceae, data=validation_heifer_fam_ADG)
summary(modelval_heifer_fam_ADG)

#Heifer FE
modelval_heifer_fam_FE <- lm(formula = FE ~ Veillonellaceae + Coriobacteriaceae + NoAssignedFamily2, data=validation_heifer_fam_FE)
summary(modelval_heifer_fam_FE)

#Validation steer data sets
validation_fam_steer <- read.table("stepwise_reg/validation_steer_fam.txt", sep="\t", header=T)
row.names(validation_fam_steer) <- validation_fam_steer[,1]
validation_fam_steer <- validation_fam_steer[,-1]
validation_steer_fam_ADFI <- merge(ADFI_res_heifer,validation_fam_steer,by="row.names")
validation_steer_fam_ADFI <- validation_steer_fam_ADFI[,-1]
colnames(validation_steer_fam_ADFI)[1] <- "ADFI"
validation_steer_fam_ADG <- merge(ADG_res_heifer,validation_fam_steer,by="row.names")
validation_steer_fam_ADG <- validation_steer_fam_ADG[,-1]
colnames(validation_steer_fam_ADG)[1] <- "ADG"
validation_steer_fam_FE <- merge(FE_res_heifer,validation_fam_steer,by="row.names")
validation_steer_fam_FE <- validation_steer_fam_FE[,-1]
colnames(validation_steer_fam_FE)[1] <- "FE"

#Validation steer models
#steer ADFI
modelval_steer_fam_ADFI <- lm(formula = ADFI ~ NoAssignedFamily13 + Paraprevotellaceae + Lachnospiraceae + Coriobacteriaceae + Bifidobacteriaceae + Erysipelotrichaceae, data=validation_steer_fam_ADFI)
summary(modelval_steer_fam_ADFI)

#steer ADG
modelval_steer_fam_ADG <- lm(formula = ADG ~ p_2534_18B5 + Lachnospiraceae + Coriobacteriaceae + Bifidobacteriaceae + NoAssignedFamily9, data=validation_steer_fam_ADG)
summary(modelval_steer_fam_ADG)

#steer FE
modelval_steer_fam_FE <- lm(formula = FE ~ NoAssignedFamily13 + Paraprevotellaceae + NoAssignedFamily9, data=validation_steer_fam_FE)
summary(modelval_steer_fam_FE)
```
